---
title: "tidyplot: An R package with diversy themes and colors for biological data visualization"
output: html_notebook
---

## Introduction

Here we developed `tidyplot` to make the data visualization easier and better. In this package, we included several commonly used themes in `ggplot2` and a large number of color combinations for biological data visualization. We'll keep updating the themes and color combinations in our package.

### 1. Installation

The dependent packages: `ggplot2` and `RColorBrewer` should be installed first.

```{r}
#install.packages("devtools")
#devtools::install_github("hmutpw/tidyplot")
library(tidyplot)
```


### 2. Instructions for colors

#### 2.1 Colors in `tidyplot` package

The `tidyplot` package includes 19 types of sequential colors with 10 different brightness. The original colors were downloaded from [material design colors](https://www.materialpalette.com/colors). Besides, we also included the colors in `RColorBrewer` package.

##### 2.1.1. To browser all the colors in `tidyplot` package:

```{r}
display.color.all(type = "material")
```

##### 2.1.2. To browser all colors in `RColorBrewer` package:

```{r}
library(RColorBrewer)
display.brewer.all()
```

##### 2.1.3. To browser combination of several beautiful colors in `tidyplot` package:

In `tidyplot` package, we designed some combinations for specific application situation, such as, embryonic stem cell development. You can directly using these
combinations to visualize your data.

```{r}
display.color.all(type = "material.set")
```

To get the color code for above combinations:

```{r}
dev_basic <- material.color.set(name = "Developmental.basic")
dev_basic
```

If you have other creative combinations, please let me known, and I'll be happy to add into out package.

#### 2.2 Get the colors from `tidyplot` package

##### 2.2.1 get colors with different brightness

If you are considering using one type of color to visualize the intensity change, e.g. heatmaps with values between 0 to 1. you can get 19 types of colors from
`tidyplot` with up to 10 different brightness. For example, you can get 6 gradient red colors:

```{r}
Reds <- material.colors(name = "Red",n = 6)
Reds
```

To visualize the above colors:

```{r}
display.color(colors = Reds, color_name = "Red")
```

##### 2.2.2 get colors from `RColorBrewer` package

Here, we can get colors from `RColorBrewer` package by using the function `brewer.colors()`:

```{r}
RdBu <- brewer.colors(name = 'RdBu',n = 11)
RdBu
```
To visualize the above colors:
```{r}
display.color(colors = RdBu, color_name = "RdBu")
```

##### 2.2.3 get specific numbers of colors

The above gradient colors may not enough to visualize heatmaps. We can use `adjust.color.number()` function to expand or shrink our color to an specific number:

```{r}
RdBu100 <- adjust.color.number(color = RdBu,n = 100)
RdBu100

```

Then, visualize them:

```{r}
display.color(colors = RdBu100, color_name = "RdBu100")
```

##### 2.3 get all colors you wanted from `tidyplot`

All the colors in `tidyplot` were deposited in `tidyplot::materialColorBrewer`, you can download the color matrix for your own design.


### 3. Instructions for themes in `ggplot2`

