---
title: 'tidyplot: An R package with diversy themes and colors for biological data
  visualization'
output:
  html_document:
    df_print: paged
---

## Introduction

Here we developed [`tidyplot`](https://github.com/hmutpw/tidyplot) to make the data visualization easier and better. In this package, we included several commonly used themes in `ggplot2` and a large number of color combinations for biological data visualization. We'll keep updating the themes and color combinations in our package.

### 1. Installation

The dependent packages:`graphics`, `ggplot2` and `RColorBrewer` should be installed first.

```{r}
#install.packages("devtools")
#devtools::install_github("hmutpw/tidyplot")
library(tidyplot)
```

### 2. Instructions for colors

#### 2.1 Colors in `tidyplot` package

The `tidyplot` package includes 19 types of sequential colors with 10 different brightness. The original colors were downloaded from [material design colors](https://www.materialpalette.com/colors). Besides, we also included the colors in `RColorBrewer` package.

##### 2.1.1. To browser all the colors in `tidyplot` package:

```{r}
display.color.all(type = "material")
```

##### 2.1.2. To browser all colors from `RColorBrewer` package:

```{r}
library(RColorBrewer)
display.brewer.all()
```

##### 2.1.3. To browser combination of several beautiful colors in `tidyplot` package:

In `tidyplot` package, we designed some combinations for specific application situation, such as, embryonic stem cell development. You can directly using these
combinations to visualize your data.

```{r}
display.color.all(type = "material.set")
```

To get the color code for above combinations:

```{r}
dev_basic <- material.color.set(name = "Developmental.basic")
dev_basic
```

If you have other creative combinations, please let me known, and I'll be happy to add into out package.

#### 2.2 Get the colors from `tidyplot` package

##### 2.2.1 get colors with different brightness

If you are considering using one type of color to visualize the intensity change, e.g. heatmaps with values between 0 to 1. you can get 19 types of colors from
`tidyplot` with up to 10 different brightness. For example, you can get 6 gradient red colors:

```{r}
Reds <- material.colors(name = "Red", brightness = 1:6)
Reds
```

To visualize the above colors:

```{r}
display.color(colors = Reds, color_name = "Red")
```

##### 2.2.2 get colors from `RColorBrewer` package

Here, we can get colors from `RColorBrewer` package by using the function `brewer.colors()`:

```{r}
RdBu <- brewer.colors(name = 'RdBu',n = 11)
RdBu
```
To visualize the above colors:
```{r}
display.color(colors = RdBu, color_name = "RdBu")
```

##### 2.2.3 get specific numbers of colors

The above gradient colors may not enough to visualize heatmaps. We can use `adjust.color.number()` function to expand or shrink our color to an specific number:

```{r}
RdBu100 <- adjust.color.number(color = RdBu,n = 100)
RdBu100

```

Then, visualize them:

```{r}
display.color(colors = RdBu100, color_name = "RdBu100")
```

#### 2.3 get all colors you wanted from `tidyplot`

All the colors in `tidyplot` were deposited in `tidyplot::materialColorBrewer`, you can download the color matrix for your own design.


### 3. Instructions for themes in `ggplot2`

Here we designed a few themes that are frequently used in data visualization.

#### 3.1 Basic theme `theme_basic()`
 
The basic theme removes background and useless panel grids. It's very useful to remove these element when you editing your figures in Adobe Illustrator (AI).

Here is the default theme:

```{r}
library(ggplot2)
library(tidyplot)
p1 <- ggplot(data = mpg, aes(x = trans, fill = factor(cyl)))+
  geom_bar(width = 0.5, color = NA)+
  labs(title = "plot title")+
  facet_wrap(.~year)
p1
```

There are few issues in default theme:

1). The title is not centered-aligned;

2). Gray background;

3). Gray strip background on each panel;

4). Gray background under legend;

5). No axis lines at x and y-axis;

Here is the updated version of theme, we modified the above issues:

1). The CNS journal requires font size between 5 - 7pt, so we set all the font size at 6pt by default;

2). The CNS journal requires line width should be larger the 0.3pt, here we set 1px by default;

3). The text color at x- and y-axis is not black, we set all the font color in black by default;

```{r}
p1+theme_basic()
```

#### 3.2 Slope theme `theme_slpoe()`

From above figure, you will find the space at x-axis is not enough to show all the items, so you can slope these items by 45 degree using `theme_slope()`.

```{r}
p1+theme_slope()
```

#### 3.3 Slope theme `theme_clean()`

If you prefer the classical theme, you can use `theme_clean()`, we removed the border and only keep the axis lines at x and y-axis:

```{r}
p1+theme_clean()
```


All above themes were summarized with my experience when perform data visualization with `ggplot2`. By removing some useless elements from
your figures will speed up editing your figures with Adobe Illustrator. You can also setting the theme.

#### 3.4 Other themes in `ggplot2`

There are lots of themes in `ggplot2`, such as `theme_bw()`, `theme_linedraw()`, `theme_light()`, `theme_dark()`, `theme_minimal()`, 
`theme_classic()`, and `theme_void()`. Please refer to [Themes](https://ggplot2-book.org/polishing.html). 

### 4. Commonly used figures in data visualization

#### 4.1 Dot plot

```{r}
library(ggplot2)
library(tidyplot)

p2 <- ggplot(data = mpg, mapping = aes(x = displ, y = cty, color = class))+
  geom_point(shape=16)+
  scale_color_manual(values = material.color.set(name = 'Discrete', brightness = 5))+
  theme_basic()

p2
```

#### 4.2 Bar plot

##### 4.2.1 Bar plot with stack position

```{r}
p3 <- ggplot(data = mpg, aes(x = trans, y = displ, fill = factor(cyl)))+
  geom_bar(stat = 'identity', position = 'stack',color=NA, width=0.75)+
  scale_fill_manual(values = material.color.set(name = 'Discrete', brightness = 5))+
  theme_slope()+
  facet_grid(.~year)

p3
```

##### 4.2.2 Bar plot with dodge position

```{r}
p4 <- ggplot(data = mpg, aes(x = trans, y = displ, fill = factor(cyl)))+
  geom_bar(stat = 'identity', position = 'dodge',color=NA)+
  scale_fill_manual(values = material.color.set(name = 'Discrete', brightness = 5))+
  theme_basic()

p4
```

##### 4.2.3 Bar plot with fill position

```{r}
p5 <- ggplot(data = mpg, aes(x = trans, y = displ, fill = factor(cyl)))+
  geom_bar(stat = 'identity', position = 'fill',color=NA)+
  scale_fill_manual(values = material.color.set(name = 'Discrete', brightness = 5))+
  theme_basic()

p5
```

##### 4.2.4 Pie plot with fill position

```{r}
p6 <- ggplot(data = mpg, aes(x = year, y = displ, fill = factor(cyl)))+
  geom_bar(stat = 'identity', position = 'fill',color=NA)+
  scale_fill_manual(values = material.color.set(name = 'Discrete', brightness = 5))+
  coord_polar(theta = "y")+
  theme_void()

p6
```

#### 4.3 Violin plot

##### 4.3.1 Violin plot

```{r}
p7 <- ggplot(data = mpg, aes(x = class, y = displ, color = class, fill=class))+
  geom_violin(color=NA)+
  geom_jitter(width=0.25, size=0.5, shape=16, color='#000000', alpha = 1)+
  scale_fill_manual(values = material.color.set(name = 'Developmental.basic', brightness = 3))+
  theme_basic()

p7
```


##### 4.3.2 Boxplot

```{r}
p8 <- ggplot(data = mpg, aes(x = class, y = displ, color = class, fill=class))+
  geom_boxplot(width=0.4,fill = NA, size=0.5,outlier.colour=NA, alpha = 0.8)+
  geom_jitter(width=0.2, size = 0.5, shape=16, color='#000000', alpha = 1)+
  scale_color_manual(values = material.color.set(name = 'Developmental.basic', brightness = 5))+
  scale_fill_manual(values = material.color.set(name = 'Developmental.basic', brightness = 3))+
  theme_basic()

p8
```

#### 4.4 Density plot

##### 4.4.1 The density distribution from the same base level

```{r}
p9 <- ggplot(data = mpg, aes(x = displ, color = class, fill=class))+
  geom_density(alpha=0.5)+
  scale_color_manual(values = material.color.set(name = 'Developmental.basic', brightness = 5))+
  scale_fill_manual(values = material.color.set(name = 'Developmental.basic', brightness = 3))+
  theme_basic()

p9
```

##### 4.4.2 Split the density distribution from y-axis

```{r}
library(ggridges)
p10 <- ggplot(data = mpg, aes(x = displ, y = class, color = class, fill=class))+
  geom_density_ridges_gradient(size = 1, scale = 1.4, panel_scaling = F, alpha=.8,na.rm=T)+
  scale_color_manual(values = material.color.set(name = 'Developmental.basic', brightness = 5))+
  scale_fill_manual(values = material.color.set(name = 'Developmental.basic', brightness = 3))+
  theme_basic()
p10
```

##### 4.4.3  Cumulative distribution plot

```{r}
p11 <- ggplot(data = mpg, aes(x = displ, color = class, fill=class))+
  stat_ecdf(geom="line")+
  scale_color_manual(values = material.color.set(name = 'Developmental.basic', brightness = 5))+
  theme_basic()
p11
```

#### 4.5. Radviz plot

##### 4.5.1  Tern plot

Tern plot describe the relative enrichment of three features, suitable for stage-specific enrichment genes among three stages.

```{r}
library(ggtern)
df <- data.frame(
  x = c(0.2, 0.5, 0.3),
  y = c(0.3, 0.2, 0.5),
  z = c(0.5, 0.3, 0.2)
)
p12 <- ggtern(data = df, mapping = aes(x = x,y = y,z = z))+
  geom_point()+
  theme_basic()

p12

```

##### 4.5.2  Radviz plot 

Here we use can visualize the relative enrichment of different types using methods in `SubLocR`, this method can used to visualize relative enrichment with more than 3 stages or any number of stages.

```{r}
devtools::install_github("hmutpw/SubLocR")
library(SubLocR)
data("ESC_RSL_data")

radviz_obj <- getRadObj(input_frm = ESC_RSL_data$D0[1:20,], anchors = c('CE','ME','SNE','CBNE','PE'))
p13 <- do_RadvizEnrichPlot(rv = radviz_obj, color_by = 'INPUT',point_size = 3)+
  geom_text(mapping = aes(label=gene_name),size=3)
p13

```


